---
title: 'Reproducible Research: Peer Assessment 1'
author: "Murat Adivar"
date: "November 13, 2015"
output: html_document
---

## Initial settings/packages
```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
```

## Loading and preprocessing the data
##### 1. Load the data (i.e. read.csv())
```{r, results='markup', warning=TRUE, message=TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
My_Data <- read.csv('activity.csv')
```
##### 2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
#My_Data$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", My_Data$interval), format='%H:%M')
```

-----

## What is mean total number of steps taken per day?
```{r}
Steps_Each_Day <- tapply(My_Data$steps, My_Data$date, sum, na.rm=TRUE)
```

##### 1. Make a histogram of the total number of steps taken each day
```{r}
qplot(Steps_Each_Day, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

##### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
Steps_Each_Day_Mean <- mean(Steps_Each_Day)
Steps_Each_Day_Median <- median(Steps_Each_Day)
```
* Mean: `r Steps_Each_Day_Mean`
* Median:  `r Steps_Each_Day_Median`

-----

## What is the average daily activity pattern?
```{r}
Avg_Steps_Per_Time_Block <- aggregate(x=list(meanSteps=My_Data$steps), by=list(interval=My_Data$interval), FUN=mean, na.rm=TRUE)
```

##### 1. Make a time series plot
```{r}
ggplot(data=Avg_Steps_Per_Time_Block, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```

##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
mostSteps <- which.max(Avg_Steps_Per_Time_Block$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", Avg_Steps_Per_Time_Block[mostSteps,'interval'])
```

* Most Steps at: `r timeMostSteps`

----

## Imputing missing values
##### 1. Calculate and report the total number of missing values in the dataset 
```{r}
NoNA <- length(which(is.na(My_Data$steps)))
```

* Number of missing values: `r NoNA`

##### 2. Devise a strategy for filling in all of the missing values in the dataset.
##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
My_Data_Imputed <- My_Data
My_Data_Imputed$steps <- impute(My_Data$steps, fun=mean)
```


##### 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r}
Steps_Each_Day_Imputed <- tapply(My_Data_Imputed$steps, My_Data_Imputed$date, sum)
qplot(Steps_Each_Day_Imputed, xlab='Total number of steps each day', ylab='Frequency using binwith 500', binwidth=500)
```


```{r}
Steps_Each_Day_Mean_Imputed <- mean(Steps_Each_Day_Imputed)
Steps_Each_Day_MedianImputed <- median(Steps_Each_Day_Imputed)
```
* Mean (Imputed): `r Steps_Each_Day_Mean_Imputed`
* Median (Imputed):  `r Steps_Each_Day_MedianImputed`


----

## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
My_Data_Imputed$dateType <-  ifelse(as.POSIXlt(My_Data_Imputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

##### 2. Make a panel plot containing a time series plot

```{r}
Avg_My_Data_Imputed <- aggregate(steps ~ interval + dateType, data=My_Data_Imputed, mean)
ggplot(Avg_My_Data_Imputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```

